### 2024-01-22 outputs/paper.qmd
Initial paper

---
title: "This is the title"
abstract: "This is the abstract"
thanks: "Code and data supporting this analysis is available at: https://github.com/prajogt/ttc_analysis.git"
author: "Timothius Prajogi"
date: today
date-format: long
format: pdf
editor: visual
toc: true
number-sections: true
bibliography: refrences.bib
---

```{r}
#| echo: false
#| message: false
#| warning: false

#### Workspace setup ####
library(janitor)
library(knitr)
library(opendatatoronto)
library(tidyverse)

```

## Introduction

## Data

Cite the packages used here.

Datasets used: subway data [@citesubwaydata], bus data [@citebusdata]

```{r}
#| echo: false
#| message: false
#| warning: false

bus_delay_data <- 
  read_csv(
     file = "data/cleaned_bus_delay_statistics.csv",
     show_col_types = FALSE
  )

subway_delay_data <-
  read_csv(
     file = "data/cleaned_subway_delay_statistics.csv",
     show_col_types = FALSE
  )
```

## Results

## Discussion

## Conclusion

## Reference