### 2024-01-23 outputs/paper.qmd
Add intro

---
title: "This is the title"
abstract: "This is the abstract"
thanks: "Code and data supporting this analysis is available at: https://github.com/prajogt/ttc_analysis.git"
author: "Timothius Prajogi"
date: today
date-format: long
format: pdf
editor: visual
toc: true
number-sections: true
bibliography: refrences.bib
---

```{r}
#| echo: false
#| message: false
#| warning: false

#### Workspace setup ####
library(janitor)
library(knitr)
library(opendatatoronto)
library(tidyverse)

```

## Introduction

The main source public transportation of public transportation in the city of 
Toronto is the TTC (Toronto Transit Commission). The most common methods of
this public transportation is subways and buses. For people who frequently
use these methods, delays are common and expected [@CityNewsBGrade]. 
So much so that leaving early is leaving on time.

With the delays getting worse and service becoming more inconsistent 
[@TorontoStarTTCDelays], it is important to analyze potential causes and 
therefore enable the TTC to explore potential solutions. One of such factors
is the tight coupling of the subway and bus systems. When issues occur with the
subway, the overflow is sent to buses, either from detoured commuter traffic,
or resource reallocation in the form of converting regular route buses to 
shuttle buses [@TorontoObserverShuttleBuses].

Within this paper we will analyze the correlations between the delays occurring
on the subway and the delays that occur for buses. In particular, the 
correlation between mechanical or security issues on the subway, which tend to 
shut down a portion of the track, and with diversion and other external 
issues (here after labeled "Miscellaneous") that occur for buses. A logical
correlation was made above, and a correlation shown in data would go to further
a cause and effect relationship.

Later we will see the general affects of the commuter load on the transit system
and how the system can be better optimized and fail-safed to prevent such frequent
delays. The paper will show the trends of the amount of delays for both
systems as a function of day and time. It was found that work days put additional
load on both systems, the total amount of delays during weekdays being 
significantly higher than on other days.

The datasets used for this analysis are described in the Data section, as well
as the process that was done to clean the data into comparable datasets.

## Data

Cite the packages used here.

Datasets used: subway data [@citesubwaydata], bus data [@citebusdata]

```{r}
#| echo: false
#| message: false
#| warning: false

# Read in cleaned data
bus_delay_data <- 
  read_csv(
     file = "data/cleaned_bus_delay_statistics.csv",
     show_col_types = FALSE
  )

subway_delay_data <-
  read_csv(
     file = "data/cleaned_subway_delay_statistics.csv",
     show_col_types = FALSE
  )
```

## Results

## Discussion

## Conclusion

## Reference